%!TEX root = ../masters_thesis.tex

\section*{Acknowledgements} % (fold)
\label{cha:acknowledgements}

The work on this Master's Thesis began in October 2015 and ended in June 2016. I want to thank my supervisors Prof. Volker Rodehorst and Prof. Sven Bertel for the continuous advices, ideas and the motivation to work on this topic of my choice.

The majority of the time of this thesis (October-December 2015 and February-May 2016) I have spent working at the \emph{Scholars' Lab} of \emph{University of Virginia}. I was provided with an excellent workspace, very cooperative people and delicious oatmeal raising cookies that all contributed to a great work atmosphere in Charlottesville, Virginia. I especially want to thank all the members of the \emph{Scholars' Lab} that were always eager to solve my problems regarding the concept and the implementation of this thesis, espacially Ammon, Eric, Scott, Purdom, Jeremy, Chris, Wayne, Ronda, Laura, Becca, Shane and Veronica. The perspective of humanity scholars on my project were very valuable for me personally and for my work. Thank you!

Since this thesis was an advancement of HistoGlobe, the project I have founded in 2010, I want to thank everybody that has ever been involved in this project. In the previous 6 years I have learned a lot about software engineering, history and geography, project management and entrepreneurship -- and how to (not) make a project successful that almost everybody likes. Special thanks goes to Frej which whom I developed the first prototype in 2011. From 2012 on, my coworkers, fellow students and friends Simon and Felix deserve great credit for everything they have done to HistoGlobe and me. Without the software they have developed until 2014 the result of this thesis would be much different. I also want to thank the other developers that have spent their free time, ideas and coding skills on the HistoGlobe project: Sascha, Sebastian, Lene, Max, Chris and Sebastian. Thank you Christian for always burning for HistoGlobe and trying to make the project profitable. Im also thankful for Carola and Katrin who have spent significant free time on HistoGlobe in 2013. Last but not least a huge Thank You to my friend and colleague Tobias for your high efforts for the HistoGlobe project, great concepts and stunning designs. I have learned from you how valuable good design is.

Neither HistoGlobe, this thesis, nor both of my academic degrees would be possible without \emph{Bauhaus-Universität Weimar} and all the associated professors and academic staff at this great instiution. Thank you for the support throughout the previous seven years. The history and the values of the Bauhaus have always inspired me and are a big part of how I think, develop and create. It has been a great time in Weimar.

Finally, I want to thank my family that has always been there for me in the first 26 years of my life. May it be my studies, my three foreign exchanges or my entrepreneurial idea -- my parents, my brother, my sister and the rest of my family have always supported me. That feeling is invaluable.

% chapter acknowledgements (end)
